\documentclass{article}

\begin{document}
\title{Mandatory assignment two, INF3200 A15\\
        Peer-to-peer leader election}
\maketitle

% arch p2p
% leader election
% joins and leaves
% minimum of 10 peers

\section{Introduction}
In this assignment we want you to implement leader election on top of a
peer-to-peer network.  All the peers in the network must agree on who the
leader is and only one peer can be leader at a time. To make this interesting
we want you to support peers leaving the network.

\section{Leader Election}
It is difficult to make everyone agree on something in a distributed system,
yet sometimes necessary. It's for eg often necessary to have one node in your
system act as leader or coordinator. Algorithms of varying complexity with
varying consistency guarantees exist, a very simple algorithm for leader
election is the bully algorithm, while a fairly complex one is paxos.

\section{Requirements}
\begin{itemize}
    \item Support at least 10 nodes in a p2p network structure of your own
    choice.

    \item Support graceful shutdown of nodes. On receiving a signal to shut
    down, a node should leave the network.

    \item Leader election. There should at all times be a single leader.
    A pertinent Q: What happens if the leader leaves the network?
\end{itemize}

\section{Hand-in}
In your report, briefly present the details of your approach to the assignment.
Then discuss the details that are interesting to you. We want to see that you
can assess the strengths and weaknesses of your implementation. Think outside
of the context of the rocks cluster.  The delivery should include:

\begin{enumerate}
    \item Source code with instructions
    on how to run.
    \item Report.
\end{enumerate}

\section{Other things}
\begin{itemize}
    \item There will be a demo presentation during the colloquium following the
    deadline. Present your work and demonstrate it.

    \item The hand in will still be on github.     

    \item Conscientious students have good experimental evaluations of their
    systems.

    \item Start early, fail early. (Make it better early.)

    \item Deadline Monday, November 2nd.
\end{itemize}

\section{References}
\begin{enumerate}
    \item \verb|https://google.com|
\end{enumerate}

\end{document}
